Learning constraints in tabular data

### 2016-05-12 Paper/ECAI/tabular.tex
minor rewriting in the introduction

In this paper,  we investigate whether machine learning techniques can be used to infer constraints (formula's and other relations) from raw spreadsheet data.
This is a new and unconventional machine learning problem.
Consider the example in Figure~\ref{fig:main_example}, where the header's names already suggest the usage of spreadsheet operations such as \textit{average} or \textit{sum}.
Looking at the first row in Table~2 and the data in Table~1, it is clear that the values in this row were computed by summing the values in the corresponding column of Table~1.
Examining Table~1 also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by taking the row-wise sum of the previous four columns.
This already provides a flavor of how this problem is different from standard data mining settings, where the data is just in rows and variables are in columns. Here everything is mixed. The data is relational on the one hand, since we have multiple table with relationships between them. And on the other hand, the data is mixed textual and numeric, since people tend to use spreadsheets for financial and accounting computations. One can see that understanding and learning constraints in tabular data is hence a new and challenging problem for machine learning. Furthermore, it is not only challenging for a machine but for humans as well.

The approach that we will introduce will borrow techniques from logical and relational learning \cite{luc_book}, where
\cite{claudien,lallouet} have studied the discovery of clausal constraints in an inductive logic programming setting, from constraint learning,  where \cite{Quacq,Conacq,modelseeker} have developed several approaches to learn conjunctive constraints in a constraint programming setting,
and from mining in databases,  where e.g. \cite{Shavlik} has applied inductive techniques to discover constraints (such as functional and multi-valued dependencies) in relational databases.  It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of function is learned in spreadsheet data from
very few examples.  Our approach contrasts with all of the mentioned approaches (except Modelseeker) in that it focusses
on learning both column- and row-constraints, it contrasts with Flashfill in learning constraints rather than functions.
A more detailed discussion on related work is contained in Section~\ref{sec:related_work}.

>>>>

In this paper,  we investigate whether learning techniques can be used to infer constraints (formula's and other relations) from raw spreadsheet data.
This is a new and unconventional machine learning problem.
Consider the example in Figure~\ref{fig:main_example}, where the header's names already suggest the usage of spreadsheet operations such as \textit{average} or \textit{sum}.
Looking at the first row in Table~2 and the data in Table~1, it is clear that the values in this row were computed by summing the values in the corresponding column of Table~1.
Examining Table~1 also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by taking the row-wise sum of the previous four columns.
This already provides a flavor of how this problem is different from standard data mining settings, where the data is just in rows and variables are in columns. Here everything is mixed. The data is relational on the one hand, since we have multiple table with relationships between them. And on the other hand, the data is mixed textual and numeric, since people tend to use spreadsheets for financial and accounting computations. One can see that understanding and learning constraints in tabular data is hence a new and challenging problem for machine learning. Furthermore, it is not only challenging for a machine but for humans as well.

The approach that we will introduce will borrow techniques from logical and relational learning \cite{luc_book}, where
the discovery of clausal constraints has been studied in an inductive logic programming setting \cite{claudien,lallouet}; and from constraint learning,  where several approaches to conjunctive constraint learning have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker}; 
and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{Shavlik}.  It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of function is learned in spreadsheet data from
very few examples.  Our approach contrasts with all of the mentioned approaches (except Modelseeker) in that it focusses
on learning both column- and row-constraints, it contrasts with Flashfill in learning constraints rather than functions.
A more detailed discussion on related work is contained in Section~\ref{sec:related_work}.
